---
title: "Dimensionality reduction for distance object describing settlement networks"
author: "Mikhail Orlov"
date: '6 ноября 2017 г '
output:
  html_document: default
pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clearance to remove possible objects etc. from previous session to avoid interference
Package loading
```{r}
rm(list = ls())

#load packages: raster, rgdal, foreach, etc.
library(rgdal)
library(raster)
library(foreach) #used in the initial script that I used
library(ggplot2)
library(RColorBrewer)
library(gridExtra)
library(factoextra)
library(FactoMineR)
library(corrplot)
library(maps)
#library(mapdata)
library(dendextend)

library(cluster)
library(caret)
#library(maxnet) #Maxent implemented in the Maxent package as well;
#But this one doesn't require additional java script and evaluated entirely in R 
#Essentially this implementation is based on glmnet algorithm
```

Loading previously saved bioclimate data on both Crimea and Vancouver island
```{r}

load('/home/jane/Документы/Misha/sheludkov/crimea_vs_vancouver_sdms/wc2.0_2.5m_bio/climate_crimea.rda')
load('/home/jane/Документы/Misha/sheludkov/crimea_vs_vancouver_sdms/wc2.0_2.5m_bio/climate_vancouver.rda')

```


To determine region of interest

```{r}

crimea <- map(ylim=c(44.3, 46), xlim=c(32.5,36.6), col='gray90', fill=TRUE)

#boundaries for which background data will be extracted
crimea_x_lims <- c(32.5, 36.6) 
crimea_y_lims <- c(44.4, 46)

abline(v = crimea_x_lims)
abline(h = crimea_y_lims)

```

Creating table of latitude and longitude for locations of interest

```{r}

Latitude_crimea <- seq(from = crimea_y_lims[1], to = crimea_y_lims[2], by = 0.062) #approx 2'
Longitude_crimea <- seq(from = crimea_x_lims[1], to = crimea_x_lims[2], by = 0.062) #approx 2'
lat_long_crimea <- expand.grid(Latitude = Latitude_crimea, Longitude = Longitude_crimea) #making pairs of latitude and longitude
Pop_crimea <- as.factor(1:nrow(lat_long_crimea)) #populations ids
pop_crimea <- cbind(lat_long_crimea, Pop_crimea) #adding Ids
row.names(pop_crimea) <- pop_crimea$Pop
#head(pop)

###NOTE: WorldClim data, even at highest res, is averaged over 1 km2.
#If your location is too close to a coast (i.e. less than 1 km2),
#there will not be any information (nothing but NAs) in this data set.
#Therefore, it may be necessary to approximate some locations by moving
#them inland. I did this using Google Earth.

#load location coordinates as SpatialPoints
for(i in pop_crimea$Pop){
  assign(i, SpatialPoints(as.matrix(t(c(pop_crimea[i,2], pop_crimea[i,1])))))
}

#check that SpatialPoints load correctly from geoTIFFs
poplist_crimea <- mget(levels(pop_crimea$Pop), envir=globalenv())

tiffvector <- unlist(list)

#Optional quality check step. For smaller datasets, will tell you which population locations should be adjusted,
#in other words, which rows are all NA. See Note above, line 51. Or check after extracting data, see line 
#foreach(p=poplist, .combine='rbind') %:%
# foreach(t=tiffvector, .combine='cbind') %do%{
#  is.na(extract(t,p))
#} #may take a while

```

Make climate data table

```{r}
# # # is a plug 

# # #climate_crimea <- foreach(p=poplist_crimea, .combine='rbind') %:%
# # #    foreach(t=tiffvector, .combine='cbind') %do%{
# # #      myValue<-extract(t, p)
# # #        } #may take a while

# # #save(climate, file = 'climate_complete_crimea_for_Orlov_Sheludkov.rda')
```


#Case region I - Crimea peninsula
##To determine region of interest
```{r}

crimea <- map(ylim=c(44.3, 46), xlim=c(32.5,36.6), col='gray90', fill=TRUE)  

#boundaries for which background data will be extracted
crimea_x_lims <- c(32.5, 36.6) 
crimea_y_lims <- c(44.4, 46)

abline(v = crimea_x_lims)
abline(h = crimea_y_lims)

#generating the set of coordinates that spans all the plotting areas 

#creating table of latitude and longitude for locations of interest
Latitude_crimea <- seq(from = crimea_y_lims[1], to = crimea_y_lims[2], by = 0.062) #approx 2'
Longitude_crimea <- seq(from = crimea_x_lims[1], to = crimea_x_lims[2], by = 0.062) #approx 2'
lat_long_crimea <- expand.grid(Latitude = Latitude_crimea, Longitude = Longitude_crimea) #making pairs of latitude and longitude

#points(x = lat_long_crimea$Longitude, y = lat_long_crimea$Latitude, pch = 2)

###Extract climate data from WorldClim.org tiles for several locations and make data table
```

Download and unzip all relevant WorldClim geoTIFF files into a single directory.
```{r}
setwd('/home/jane/Документы/Misha/sheludkov/crimea_vs_vancouver_sdms/wc2.0_2.5m_bio/')
#check the folder content
dir()
```


Read names of all .tif (geotif) files in directory into a list
#taken from http://stackoverflow.com/questions/5319839/read-multiple-csv-files-into-separate-data-frames

```{r}
filenames <- grep('*.tif', list.files(), value = T)

#Load all geoTIFF files
for(i in filenames){
  filepath <- file.path(i)
  assign(i, raster(filepath))
}

#check that all files loaded properly by raster
#taken from http://stackoverflow.com/questions/15387727/use-object-names-as-list-names-in-r
list <- mget(filenames, envir=globalenv())

for(i in list){
  if (hasValues(i)==FALSE){
    print(i,"hasValues error")
  }
  if (inMemory(i)==TRUE){
    print(i, "inMemory error")
  }
  else{
    print("All checked out!")
  }
}

#Vacouver island data extraction
Pop_crimea <- as.factor((1:nrow(lat_long_crimea))) #populations ids
pop_crimea <- cbind(lat_long_crimea, Pop_crimea) #adding Ids
row.names(pop_crimea) <- pop_crimea$Pop
#head(pop_crimea)

###NOTE: WorldClim data, even at highest res, is averaged over 1 km2.
#If your location is too close to a coast (i.e. less than 1 km2),
#there will not be any information (nothing but NAs) in this data set.
#Therefore, it may be necessary to approximate some locations by moving
#them inland. I did this using Google Earth.

#load location coordinates as SpatialPoints
for(i in pop_crimea$Pop){
  assign(i, SpatialPoints(as.matrix(t(c(pop_crimea[i,2], pop_crimea[i,1])))))
}

#check that SpatialPoints load correctly from geoTIFFs
poplist_crimea <- mget(levels(pop_crimea$Pop), envir=globalenv())

tiffvector <- unlist(list)

#Optional quality check step. For smaller datasets, will tell you which population locations should be adjusted,
#in other words, which rows are all NA. See Note above, line 51. Or check after extracting data, see line 
#foreach(p=poplist, .combine='rbind') %:%
# foreach(t=tiffvector, .combine='cbind') %do%{
#  is.na(extract(t,p))
#} #may take a while
```

###make climate data table
```{r}

# # # is a plug 

# # #climate <- foreach(p=poplist, .combine='rbind') %:%
# # #  foreach(t=tiffvector, .combine='cbind') %do%{
# # #    myValue<-extract(t, p)
# # #    } #may take a while

# # #save(climate, file = 'climate_complete_crimea_for_Orlov_Sheludkov.rda')
load('/home/jane/Документы/Misha/sheludkov/crimea_vs_vancouver_sdms/wc2.0_2.5m_bio/climate_crimea.rda')
```

#Case region II - Vancouver island
##To determine region of interest
```{r}

#boundaries for which background data will be extracted
vancouver_x_lims <- c(-129, -123) 
vancouver_y_lims <- c(48, 51)

vancouver <- map(ylim=vancouver_y_lims, xlim=vancouver_x_lims, col='gray90', fill=TRUE)  

#building polygon by vertices
xs <- min(vancouver_x_lims) + c(5.6, 5.5, 5.1, 3.9, 3.5, 2.2, 0.5, 0.8, 3.4, 5.4, 5.8)
ys <- min(vancouver_y_lims) + c(0.7, 0.7, 1.2, 1.95, 2.35, 2.6, 3, 1.9, 0.8, 0.2, 0.5)

for (i in seq_along(xs)){
  points(x = xs[i], y = ys[i], pch = 10)
}
#adding the polygon to define the area for which data will be extracted
polygon(x = xs, y = ys, col = 'grey90')

#generating the set of coordinates that spans all the plotting areas 

#creating table of latitude and longitude for locations of interest
Latitude_vancouver <- seq(from = vancouver_y_lims[1], to = vancouver_y_lims[2], by = 0.062) #approx 2'
Longitude_vancouver <- seq(from = vancouver_x_lims[1], to = vancouver_x_lims[2], by = 0.062) #approx 2'
lat_long_vancouver <- expand.grid(Latitude = Latitude_vancouver, Longitude = Longitude_vancouver) #making pairs of latitude and longitude

#points(x = lat_long_vancouver$Longitude, y = lat_long_vancouver$Latitude, pch = 2)

abline(v = vancouver_y_lims)
abline(h = vancouver_x_lims)
```


Which of these fall within the polygon?
Data for Vancouver hold points outside the island; those needs to be deleted
```{r}

inds <- which(point.in.polygon(point.x = lat_long_vancouver$Longitude, point.y = lat_long_vancouver$Latitude, pol.x = xs, pol.y = ys)==1)

coords_in_polygons <- lat_long_vancouver[inds, ]

climate_vancouver_island_only <- climate[inds,]


#View(climate_vancouver_island_only)
```

Extract climate data from WorldClim.org tiles for several locations and make data table.
Ddownload and unzip all relevant WorldClim geoTIFF files into a single directory.

```{r}

setwd('/home/jane/Документы/Misha/sheludkov/crimea_vs_vancouver_sdms/wc2.0_2.5m_bio/')
#check the folder content
dir()


#Read names of all .tif (geotif) files in directory into a list
#taken from http://stackoverflow.com/questions/5319839/read-multiple-csv-files-into-separate-data-frames
filenames <- grep('*.tif', list.files(), value = T)

#Load all geoTIFF files
for(i in filenames){
  filepath <- file.path(i)
  assign(i, raster(filepath))
}

#check that all files loaded properly by raster
#taken from http://stackoverflow.com/questions/15387727/use-object-names-as-list-names-in-r
list <- mget(filenames, envir=globalenv())

for(i in list){
  if (hasValues(i)==FALSE){
    print(i,"hasValues error")
  }
  if (inMemory(i)==TRUE){
    print(i, "inMemory error")
  }
  else{
    print("All checked out!")
  }
}

#Vacouver island data extraction
Pop_vancouver <- as.factor((1:nrow(lat_long_vancouver))[inds]) #populations ids
pop_vancouver <- cbind(lat_long_vancouver[inds,], Pop_vancouver) #adding Ids
row.names(pop_vancouver) <- pop_vancouver$Pop
#head(pop_vancouver)

###NOTE: WorldClim data, even at highest res, is averaged over 1 km2.
#If your location is too close to a coast (i.e. less than 1 km2),
#there will not be any information (nothing but NAs) in this data set.
#Therefore, it may be necessary to approximate some locations by moving
#them inland. I did this using Google Earth.

#load location coordinates as SpatialPoints
for(i in pop_vancouver$Pop){
  assign(i, SpatialPoints(as.matrix(t(c(pop_vancouver[i,2], pop_vancouver[i,1])))))
}

#check that SpatialPoints load correctly from geoTIFFs
poplist_vancouver <- mget(levels(pop_vancouver$Pop), envir=globalenv())

tiffvector <- unlist(list)

#Optional quality check step. For smaller datasets, will tell you which population locations should be adjusted,
#in other words, which rows are all NA. See Note above, line 51. Or check after extracting data, see line 
#foreach(p=poplist, .combine='rbind') %:%
# foreach(t=tiffvector, .combine='cbind') %do%{
#  is.na(extract(t,p))
#} #may take a while

```


###make climate data table
```{r}

# # # is a plug 

# # #climate <- foreach(p=poplist, .combine='rbind') %:%
# # #  foreach(t=tiffvector, .combine='cbind') %do%{
# # #    myValue<-extract(t, p)
# # #    } #may take a while

# # #save(climate, file = 'climate_complete_crimea_for_Orlov_Sheludkov.rda')


#tidying table
popnames_crimea <- sort(as.character(pop_crimea$Pop))
clim_crimea <- as.data.frame(climate_crimea, row.names=popnames_crimea)

#check for NAs
movepops_crimea <- clim_crimea[rowSums(is.na(clim_crimea)) == ncol(clim_crimea),] #from the initial script from the Internet
#or
table(complete.cases(clim_crimea))['TRUE'] #rows that are not NAs

head(clim_crimea)
# # #View(clim)
#how many points there are in total
unique(lapply(clim_crimea, length))

#matrix to data frame transformation; probably there's a better way
res_crimea <- c()
for (i in clim_crimea) {
  res_crimea <- cbind(res_crimea, i)
}

#setting bioclimatic variable names
raw <- 'BIO1 = Annual Mean Temperature, BIO2 = Mean Diurnal Range (Mean of monthly (max temp - min temp)), BIO3 = Isothermality (BIO2/BIO7) (* 100), BIO4 = Temperature Seasonality (standard deviation *100), BIO5 = Max Temperature of Warmest Month, BIO6 = Min Temperature of Coldest Month, BIO7 = Temperature Annual Range (BIO5-BIO6), BIO8 = Mean Temperature of Wettest Quarter, BIO9 = Mean Temperature of Driest Quarter, BIO10 = Mean Temperature of Warmest Quarter, BIO11 = Mean Temperature of Coldest Quarter, BIO12 = Annual Precipitation, BIO13 = Precipitation of Wettest Month, BIO14 = Precipitation of Driest Month, BIO15 = Precipitation Seasonality (Coefficient of Variation), BIO16 = Precipitation of Wettest Quarter, BIO17 = Precipitation of Driest Quarter, BIO18 = Precipitation of Warmest Quarter, BIO19 = Precipitation of Coldest Quarter'
#raw1 <- gsub('BIO.*? = ', '', raw)
bioclim_vars <- unlist(strsplit(raw, split = ', '))
bioclim_vars <- substr(bioclim_vars, start = 8, 100)

#to restore order
rownames(res_crimea) <- as.numeric(popnames_crimea)
res_crimea <- cbind(pop_crimea, res_crimea) #population Ids are added
colnames(res_crimea)[4:22] <- bioclim_vars

#removing NAs
res_no_nas_crimea <- res_crimea[complete.cases(res_crimea),]
res_no_nas_scaled_crimea <- as.data.frame(scale(res_no_nas_crimea[,-c(1:3)], center = T, scale = T))
initial_res_no_nas_scaled_crimea <- res_no_nas_scaled_crimea
# # #save(initial_res_no_nas_scaled, file = 'initial_complete_climate_no_nas_scaled.Rda')
load('/home/jane/Документы/Misha/sheludkov/crimea_vs_vancouver_sdms/wc2.0_2.5m_bio/climate_vancouver.rda')
# # #now to retain information on the island only

climate_vancouver <- climate[inds, ]
popnames_vancouver <- row.names(pop_vancouver)[inds]
str(climate_vancouver)
#how many NAs are they?
table(complete.cases(climate))

# # #
dim(climate_vancouver)
# # #
class(climate_vancouver)

#tidying table
popnames_vancouver <- sort(as.character(pop_vancouver$Pop))[inds]
clim_vancouver <- as.data.frame(climate_vancouver, row.names=popnames_vancouver)

#check for NAs
movepops <- clim_vancouver[rowSums(is.na(clim_vancouver)) == ncol(clim_vancouver),] #from the initial script from the Internet
#or
table(complete.cases(clim_vancouver))['TRUE'] #rows that are not NAs

head(clim_vancouver)
# # #View(clim)
#how many points there are in total
unique(lapply(clim_vancouver, length))

#matrix to data frame transformation; probably there's a better way
res_vancouver <- c()
for (i in clim_vancouver) {
  res_vancouver <- cbind(res_vancouver, i)
}

#setting bioclimatic variable names
raw <- 'BIO1 = Annual Mean Temperature, BIO2 = Mean Diurnal Range (Mean of monthly (max temp - min temp)), BIO3 = Isothermality (BIO2/BIO7) (* 100), BIO4 = Temperature Seasonality (standard deviation *100), BIO5 = Max Temperature of Warmest Month, BIO6 = Min Temperature of Coldest Month, BIO7 = Temperature Annual Range (BIO5-BIO6), BIO8 = Mean Temperature of Wettest Quarter, BIO9 = Mean Temperature of Driest Quarter, BIO10 = Mean Temperature of Warmest Quarter, BIO11 = Mean Temperature of Coldest Quarter, BIO12 = Annual Precipitation, BIO13 = Precipitation of Wettest Month, BIO14 = Precipitation of Driest Month, BIO15 = Precipitation Seasonality (Coefficient of Variation), BIO16 = Precipitation of Wettest Quarter, BIO17 = Precipitation of Driest Quarter, BIO18 = Precipitation of Warmest Quarter, BIO19 = Precipitation of Coldest Quarter'
#raw1 <- gsub('BIO.*? = ', '', raw)
bioclim_vars <- unlist(strsplit(raw, split = ', '))
bioclim_vars <- substr(bioclim_vars, start = 8, 100)

#to restore order
rownames(res_vancouver) <- as.numeric(popnames_vancouver)
res_vancouver <- cbind(pop_vancouver, res_vancouver) #population Ids are added
colnames(res_vancouver)[4:22] <- bioclim_vars

#removing NAs
res_no_nas_vancouver <- res_vancouver[complete.cases(res_vancouver),]
res_no_nas_scaled_vancouver <- as.data.frame(scale(res_no_nas_vancouver[,-c(1:3)], center = T, scale = T))
initial_res_no_nas_scaled_vancouver <- res_no_nas_scaled_vancouver
```
#Visualization part

Check up first
```{r}
dim(res_no_nas_crimea); dim(res_no_nas_vancouver)
# # #View(res)
#plots just to check the data

#visualization of certain variables (colouring by isothermality)
#vector by which the color will be set
#vectocol <- (min(sort(unique(na.omit(res_no_nas_crimeas$`Isothermality (BIO2/BIO7) (* 100)`)))):max(sort(unique(na.omit(res_no_nas_crimea$`Isothermality (BIO2/BIO7) (* 100)`)))))
```

```{r Isothermality comparisom}


ggplot_Isothermality_crimea <- ggplot(data = res_no_nas_scaled_crimea, mapping = aes(x = res_no_nas_crimea$Longitude, #there is no lattitude and longitude in scaled df
                                                                              y = res_no_nas_crimea$Latitude, 
                                                                              lwd = 1, alpha = res_no_nas_scaled_crimea$`Isothermality (BIO2/BIO7) (* 100)`, 
                                                                              color = res_no_nas_scaled_crimea$`Isothermality (BIO2/BIO7) (* 100)`))+geom_point(show.legend = F) + ggtitle(label = 'Crimea, Isothermality (BIO 2/BIO 7)')+coord_fixed()

ggplot_Isothermality_vancouver <- ggplot(data = res_no_nas_scaled_vancouver, mapping = aes(x = res_no_nas_vancouver$Longitude, 
                                                                                    y = res_no_nas_vancouver$Latitude, 
                                                                                    lwd = 1, alpha = res_no_nas_scaled_vancouver$`Isothermality (BIO2/BIO7) (* 100)`, 
                                                                                    color = res_no_nas_scaled_vancouver$`Isothermality (BIO2/BIO7) (* 100)`))+geom_point(show.legend = F) + ggtitle(label = 'Vancouver, Isothermality (BIO 2/BIO 7)')+coord_fixed()

grid.arrange(ggplot_Isothermality_crimea, ggplot_Isothermality_vancouver, ncol = 2)
```

```{r Precipitation of Wettest Month comparisom}
ggplot_Precipitation_of_Wettest_Month_crimea <- ggplot(data = res_no_nas_crimea, mapping = aes(x = res_no_nas_crimea$Longitude, 
                                                                              y = res_no_nas_crimea$Latitude, 
                                                                              lwd = 1, alpha = res_no_nas_scaled_crimea$` Precipitation of Wettest Month`, 
                                                                              color = res_no_nas_scaled_crimea$` Precipitation of Wettest Month`))+geom_point(show.legend = F) + ggtitle(label = 'Crimea, Precipitation of Wettest Month')+coord_fixed()

ggplot_Precipitation_of_Wettest_Month_vancouver <- ggplot(data = res_no_nas_vancouver, mapping = aes(x = res_no_nas_vancouver$Longitude, 
                                                                                    y = res_no_nas_vancouver$Latitude, 
                                                                                    lwd = 1, alpha = res_no_nas_scaled_vancouver$`Isothermality (BIO2/BIO7) (* 100)`, 
                                                                                    color = res_no_nas_scaled_vancouver$`Isothermality (BIO2/BIO7) (* 100)`))+geom_point(show.legend = F) + ggtitle(label = 'Vancouver,Precipitation of Wettest Month')+coord_fixed()

grid.arrange(ggplot_Precipitation_of_Wettest_Month_crimea, ggplot_Precipitation_of_Wettest_Month_vancouver, ncol = 2)
```

```{r Seasonality comparisom}

ggplot_Seasonality_crimea <- ggplot(data = res_no_nas_crimea, mapping = aes(x = res_no_nas_crimea$Longitude, 
                                                                                               y = res_no_nas_crimea$Latitude, lwd = 1, alpha = res_no_nas_scaled_crimea$`Temperature Seasonality (standard deviation *100)`, 
                                                                                               color = res_no_nas_scaled_crimea$`Temperature Seasonality (standard deviation *100)`))+geom_point(show.legend = F) + ggtitle(label = 'Crimea, Seasonality')+coord_fixed()

ggplot_Seasonality_vancouver <- ggplot(data = res_no_nas_vancouver, mapping = aes(x = res_no_nas_vancouver$Longitude, 
                                                                                                     y = res_no_nas_vancouver$Latitude, 
                                                                                                     lwd = 1, alpha = res_no_nas_scaled_vancouver$`Temperature Seasonality (standard deviation *100)`, 
                                                                                                     color = res_no_nas_scaled_vancouver$`Temperature Seasonality (standard deviation *100)`))+geom_point(show.legend = F) + ggtitle(label = 'Vancouver, Seasonality')+coord_fixed()

grid.arrange(ggplot_Seasonality_crimea, ggplot_Seasonality_vancouver, ncol = 2)
```


#Exploratory analysis

```{r correlation analysis (Pearson correlation coefficient)}

#setting shorter names
colnames(res_no_nas_crimea) <- paste('BIO', -2:19)
colnames(res_no_nas_vancouver) <- paste('BIO', -2:19) 
# # #svg('initial_corrplot.svg', height = 6, width = 6)
#corrplot(cor((res_no_nas_crimea[,-c(1:3)]), method = 'p'), col= colorRampPalette(c("grey75", "grey15"))(10), tl.col = 1)

par(mfrow = c(1, 2))
corrplot(cor((res_no_nas_crimea[,-c(1:3)]), method = 'p'), tl.col = 1)
corrplot(cor((res_no_nas_vancouver[,-c(1:3)]), method = 'p'), tl.col = 1)
# # #dev.off()
```


```{r PCA (principal component analysis), base R}

#Compute PCA in R using prcomp()
res.pca_crimea <- prcomp(res_no_nas_crimea[,-c(1:3)], scale = TRUE) #no columns corresponding to coordinates and population names
res.pca_vancouver <- prcomp(res_no_nas_vancouver[,-c(1:3)], scale = TRUE) #no columns corresponding to coordinates and population names

#Visualize eigenvalues (scree plot). Show the percentage of variances explained by each principal component.
ggplot_screeplot_crimea <- fviz_eig(res.pca_crimea) + ylim(c(0, 60))
ggplot_screeplot_vancouver <- fviz_eig(res.pca_vancouver) + ylim(c(0, 60))

grid.arrange(ggplot_screeplot_crimea, ggplot_screeplot_vancouver, ncol = 2)

```

```{r Biplots}

biplot_crimea <- fviz_pca_biplot(res.pca_crimea, repel = TRUE,
                                 col.var = "contrib", # Variables color
                                 col.ind = 'grey75', 
                                 fill.ind = 'grey75',
                                 label = 'var',
                                 alpha_ind = 2# Individuals color
)+  scale_colour_gradient(low = "grey10", high = "black") +
  theme(legend.position = 'none')

biplot_vancouver <- fviz_pca_biplot(res.pca_vancouver, repel = TRUE,
                                 col.var = "contrib", # Variables color
                                 col.ind = 'grey75', 
                                 fill.ind = 'grey75',
                                 label = 'var',
                                 alpha_ind = 2# Individuals color
)+  scale_colour_gradient(low = "grey10", high = "black") +
  theme(legend.position = 'none')

grid.arrange(biplot_crimea, biplot_vancouver, ncol = 2)
```

```{r Hierarchical clusterization etc.}


hclusted_crimea <- hclust(dist(res_no_nas_scaled_crimea), method = 'ward.D2')
hclusted_vancouver <- hclust(dist(res_no_nas_scaled_vancouver), method = 'ward.D2')

# # #svg('initial_hclust_no_numbers.svg', height = 4, width = 6)
par(mfrow = c(1, 2),
    mar = rep(3, 4))
plot(hclusted_crimea, labels = F, main = "Ward's method clusterization for\n complete bioclimate data for Crimea", xaxt = 'n', xlab = '', col = 'grey60', sub ='', ylim = c(-10, 100))
rect.hclust(hclusted_crimea, k = 5, border =  1)
text(x = c(25, 125, 305, 440, 825), y = rep(-15,5),  as.roman(1:5), cex = 0.9)

plot(hclusted_vancouver, labels = F, main = "Ward's method clusterization for\n complete bioclimate data for Crimea", xaxt = 'n', xlab = '', col = 'grey60', sub ='', ylim = c(-10, 100))
rect.hclust(hclusted_vancouver, k = 4, border =  1)
text(x = c(25, 125, 305, 440, 825), y = rep(-15,5),  as.roman(1:5), cex = 0.9)
# # #dev.off()
```

```{r Cutting hclust object, eval=F}

cut_hclusted_crimea <- cutree(hclusted_crimea, k = 5)
cut_hclusted_vancouver <- cutree(hclusted_vancouver, k = 4)


map_bioclimate_crimea <- ggplot(data = res_no_nas_crimea, mapping = aes(y = res_no_nas_crimea$Latitude, 
                                                          x = res_no_nas_crimea$Longitude, 
                                                          lwd = 0.5, 
                                                          color = rainbow(length(unique(cut_hclusted_crimea)))[cut_hclusted_crimea]))+
  geom_point(show.legend = F) + ylab('Latitude') + xlab('Longitude') + 
ggtitle(label = "Generic map") + coord_fixed()

map_bioclimate_vancouver <- ggplot(data = res_no_nas_vancouver, mapping = aes(y = res_no_nas_vancouver$Latitude, 
                                                          x = res_no_nas_vancouver$Longitude, 
                                                          lwd = 0.5, 
                                                          color = rainbow(length(unique(cut_hclusted_vancouver)))[cut_hclusted_vancouver]))+
  geom_point(show.legend = F) + ylab('Latitude') + xlab('Longitude') + 
ggtitle(label = "Generic map") + coord_fixed()

grid.arrange(map_bioclimate_crimea, map_bioclimate_vancouver, ncol = 2)

```